Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-10-05 SupportingInformation.Rmd
adding modelsel and npa rtesults for all aphid sp : firstflight

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
\caption[Testing non-parametric statistics for Ceder-creek data with $numbin=10$ for year 2000]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{birds' BMR data}. As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).  When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than. \label{fig_bmr_birds_nonparam}}
\end{center}
\end{figure}

```{r tab_bmr_birds_asym, echo=F, results="markup"}
tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
knitr::kable(tab_bmr_birds_asym,
             format="pandoc",
             caption = "**Birds' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis on year 2000 \\label{tab_bmr_birds_asym}",
             booktabs=T)
```

@@

set.seed(seed)
library(parallel)
source("FittingCopula_selective_loc.R")
RES_all_sp_aphid_count<-mclapply(X=c(1:20),FUN=RES_single_sp,d_allsp=data_aphid_count,
                                 families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=35,
                                 good_loc_given=F,good_loc=NA,mc.cores=3)
saveRDS(RES_all_sp_aphid_count,"./Results/fitting_results/RES_all_sp_aphid_count.RDS")

@@

for(sp in c(1:20)){
  #cat("sp=",sp,"\n")
  summary_aphid_count_allsp$species[sp]<-sp
  LTmUT<-(RES_all_sp_aphid_count[[sp]]$LTdep_AICw - RES_all_sp_aphid_count[[sp]]$UTdep_AICw) # total loc pair 182
  summary_aphid_count_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
  summary_aphid_count_allsp$`#LTmUT<0`[sp]<-sum(LTmUT<0,na.rm=T)
  ind_LTmUT_pos<-which(LTmUT>0,arr.ind = T)
  summary_aphid_count_allsp$`val_LTmUT>0`[sp]<-sum(LTmUT[ind_LTmUT_pos])
  ind_LTmUT_neg<-which(LTmUT<0,arr.ind = T)
  summary_aphid_count_allsp$`val_LTmUT<0`[sp]<-sum(LTmUT[ind_LTmUT_neg])

@@

@@

}
#----------------------------------
#calling the above function on bivariable datasets
m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke,rs=stat_soilCN_ga)
m2<-bivar_tab_summary_fn(r=BivMS_birds,rs=stat_birds_ga)
m3<-bivar_tab_summary_fn(r=BivMS_res_cc$`2000`,rs=stat_cc_ga$`2000`)
tab_bivar_summary<-cbind(m1,m2,m3)
colnames(tab_bivar_summary)<-c("SoilCN","Birds' BMR","Ceder-creek")
saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
```

>>>>

```{r read_BMR_mammals_and_copulaplot, echo=F, results='hide'}
set.seed(seed)
raw_m<-readxl::read_xls("./Data/BMR/MammalianBodyMassesMetabolicRates/MammalianMassesAndMetabolicRatesMcNab2008.xls")
raw_m<-raw_m[,1:3]
raw_m<-na.omit(raw_m)
raw_m<-raw_m[,2:3]

m<-log10(raw_m)
colnames(m)<-c("log10M(g)","log10BMR(KJ/h)")
m<-as.data.frame(m)

pdf("./Results/BMR_results/mammals/raw_mammals_BMR.pdf")
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(m[,1],m[,2],type="p",col="red",xlab="log(Mass)",ylab="log(BMR)",cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()

source("getcopula.R")
pdf("./Results/BMR_results/mammals/copula_mammals_BMR.pdf")
cop_m<-getcopula(d=m,rankon=T,ploton=T) #convert to a copula and plot
dev.off()
```

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
\caption[Testing non-parametric statistics for birds' BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{birds' BMR data}. As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).  When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than. \label{fig_bmr_birds_nonparam}}
\end{center}
\end{figure}

```{r tab_bmr_birds_asym, echo=F, results="markup"}
tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
knitr::kable(tab_bmr_birds_asym,
             format="pandoc",
             caption = "**Birds' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_bmr_birds_asym}",
             booktabs=T)
```

```{r bivmodselect_mammals_BMR, echo=F, cache=T, cache.extra=list(seed,cop_m,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
BivMS_mammals<-OurBiCopSelect(u1=cop_m[,1],u2=cop_m[,2],families=c(1,3:10,13:14,16:20),
                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                              gofnormal=F,status=F)
saveRDS(BivMS_mammals,file="./Results/BMR_results/mammals/BivMS_mammals.RDS")
```

<!--Display model selection results for BMR:birds data in a table-->
```{r tab_bmr_mammals_fit, echo=F, results='markup'}
h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for **mammals' BMR dataset** . \\label{tab_bmr_birds_fit}",
             booktabs=T)
```

```{r stat_bmr_mammals_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_m,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)
stat_mammals_ga<-bivfunctionplot(v=cop_m,resloc="./Results/BMR_results/mammals/",nametag="mammals_numbin_10",numbin = 10)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
\caption[Testing non-parametric statistics for mammals' BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{mammals' BMR data}. As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).  When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than. \label{fig_bmr_mammals_nonparam}}
\end{center}
\end{figure}

```{r tab_bmr_mammals_asym, echo=F, results="markup"}
tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
knitr::kable(tab_bmr_mammals_asym,
             format="pandoc",
             caption = "**Mammals' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_bmr_mammals_asym}",
             booktabs=T)
```

@@

set.seed(seed)
library(parallel)
source("FittingCopula_selective_loc.R")
RES_all_sp_aphid_count<-mclapply(X=c(1:20),FUN=RES_single_sp,d_allsp=data_aphid_count,
                                 families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                                 good_loc_given=F,good_loc=NA,mc.cores=3)
saveRDS(RES_all_sp_aphid_count,"./Results/fitting_results/RES_all_sp_aphid_count.RDS")

@@

for(sp in c(1:20)){
  #cat("sp=",sp,"\n")
  summary_aphid_count_allsp$species[sp]<-sp
  LTmUT<-(RES_all_sp_aphid_count[[sp]]$LTdep_AICw - RES_all_sp_aphid_count[[sp]]$UTdep_AICw) 
  summary_aphid_count_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
  summary_aphid_count_allsp$`#LTmUT<0`[sp]<-sum(LTmUT<0,na.rm=T)
  ind_LTmUT_pos<-which(LTmUT>0,arr.ind = T)
  summary_aphid_count_allsp$`val_LTmUT>0`[sp]<-sum(LTmUT[ind_LTmUT_pos])
  ind_LTmUT_neg<-which(LTmUT<0,arr.ind = T)
  summary_aphid_count_allsp$`val_LTmUT<0`[sp]<-sum(LTmUT[ind_LTmUT_neg])

@@

```{r modelsel_and_npa_aphid_ff_allsp, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

set.seed(seed)
library(parallel)
source("FittingCopula_selective_loc.R")
RES_all_sp_aphid_ff<-mclapply(X=c(1:20),FUN=RES_single_sp,d_allsp=data_aphid_ff,
                                 families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                                 good_loc_given=F,good_loc=NA,mc.cores=3)
saveRDS(RES_all_sp_aphid_ff,"./Results/fitting_results/RES_all_sp_aphid_ff.RDS")

source("good_loclist.R")
source("NonParamStat.R")
resloc<-"./Results/stat_results/stat_aphid_ff/"
longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)

stat_aphid_ff_allsp<-vector("list",20)

for(sp in 1:20){
  nm<-paste("sp_",sp,sep="")
  resloc2<-paste(resloc,nm,sep="")
   if (!dir.exists(resloc2)){
    dir.create(resloc2)
  }
  good_loc<-good_loclist(d_allsp=data_aphid_ff,sp=sp,data_pt_thrs=30)

  temp<-multcall(d_allsp=data_aphid_ff,sp=sp,lats=lats,longs=longs,
                 pfname=paste(resloc2,"/","Sp_",sp,sep=''),good_loc=good_loc)
  stat_aphid_ff_allsp[[sp]]<-temp
}
saveRDS(stat_aphid_ff_allsp,paste(resloc,file="stat_aphid_ff_all_sp.RDS",sep=''))

summary_aphid_ff_allsp<-as.data.frame(matrix(NA,nrow=20,ncol=8))
colnames(summary_aphid_ff_allsp)<-c("species","#LTmUT>0","#LTmUT<0","val_LTmUT>0","val_LTmUT<0",
                                     "CorlmCoru_mean_CI","PlmPu_mean_CI","D2umD2l_mean_CI")

for(sp in c(1:20)){
  #cat("sp=",sp,"\n")
  summary_aphid_ff_allsp$species[sp]<-sp
  LTmUT<-(RES_all_sp_aphid_ff[[sp]]$LTdep_AICw - RES_all_sp_aphid_ff[[sp]]$UTdep_AICw) 
  summary_aphid_ff_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
  summary_aphid_ff_allsp$`#LTmUT<0`[sp]<-sum(LTmUT<0,na.rm=T)
  ind_LTmUT_pos<-which(LTmUT>0,arr.ind = T)
  summary_aphid_ff_allsp$`val_LTmUT>0`[sp]<-sum(LTmUT[ind_LTmUT_pos])
  ind_LTmUT_neg<-which(LTmUT<0,arr.ind = T)
  summary_aphid_ff_allsp$`val_LTmUT<0`[sp]<-sum(LTmUT[ind_LTmUT_neg])

  summary_aphid_ff_allsp$CorlmCoru_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_ff_allsp[[sp]]$numericdf[9,c(2:4)],3))))
  summary_aphid_ff_allsp$PlmPu_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_ff_allsp[[sp]]$numericdf[10,c(2:4)],3))))
  summary_aphid_ff_allsp$D2umD2l_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_ff_allsp[[sp]]$numericdf[11,c(2:4)],3))))
}
saveRDS(summary_aphid_ff_allsp,"./Results/summary_aphid_ff_allsp.RDS")
```

@@

}
#----------------------------------
#calling the above function on bivariable datasets
m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke,rs=stat_soilCN_ga)
m2<-bivar_tab_summary_fn(r=BivMS_birds,rs=stat_birds_ga)
m3<-bivar_tab_summary_fn(r=BivMS_mammals,rs=stat_mammals_ga)
m4<-bivar_tab_summary_fn(r=BivMS_res_cc$`2000`,rs=stat_cc_ga$`2000`)
tab_bivar_summary<-cbind(m1,m2,m3,m4)
colnames(tab_bivar_summary)<-c("SoilCN","Birds' BMR","Mammals' BMR","Ceder-creek")
saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
```
